

# Global Plastic Waste  

```{r,message=F}
library(tidyverse)
library(janitor)
theme_set(theme_light())
knitr::opts_chunk$set(message=F)
```


```{r}
# loading
coast_vs_waste <- read_csv("data/global_plastic_waste/coastal-population-vs-mismanaged-plastic.csv")
mismanaged_vs_gdp <- read_csv("data/global_plastic_waste/per-capita-mismanaged-plastic-waste-vs-gdp-per-capita.csv")
waste_vs_gdp <- read_csv("data/global_plastic_waste/per-capita-plastic-waste-vs-gdp-per-capita.csv")
```


Take a glance at the three datasets:   

```{r}
coast_vs_waste
mismanaged_vs_gdp
waste_vs_gdp
```

**Data dictionary**  


`coast_vs_waste.csv`

|variable             |class     |description |
|:---|:---|:-----------|
|Entity | Character | Country Name |
| Code | Character | 3 Letter country code |
| Year | Integer (date) | Year |
| Mismanaged plastic waste (tonnes) | double | Tonnes of mismanaged plastic waste |
| Coastal population | Double | Number of individuals living on/near coast |
| Total Population | double | Total population according to Gapminder |


 `mismanaged_vs_gdp.csv`

|variable             |class     |description |
|:---|:---|:-----------|
|Entity | Character | Country Name |
| Code | Character | 3 Letter country code |
| Year | Integer (date) | Year |
| Per capita mismanaged plastic waste (kg per day) | double | Amount of mismanaged plastic waste per capita in kg/day |
| GDP per capita | Double | GDP per capita constant 2011 international $, rate  |
| Total Population | double | Total population according to Gapminder |

 `waste_vs_gdp.csv`

|variable             |class     |description |
|:---|:---|:-----------|
|Entity | Character | Country Name |
| Code | Character | 3 Letter country code |
| Year | Integer (date) | Year |
| Per capita plastic waste (kg per person per day) | double | Amount of  plastic waste per capita in kg/day |
| GDP per capita | Double | GDP per capita constant 2011 international $, rate  |
| Total Population | double | Total population according to Gapminder |


We see from all these 3 datasets that:  

* bad variable naming: use `janitor::clean_names` and then `rename` (for some names are still too long or hard to understand)  

* large fraction of `NA`s:  

```{r}
coast_vs_waste %>%
  summarize_all(~ mean(is.na(.)))
```

This is because they actually all contains data only from year 2010:  

```{r}
coast_vs_waste %>%
  filter(!is.na(`Mismanaged plastic waste (tonnes)`)) %>%
  count(Year)
```


```{r}
mismanaged_vs_gdp %>% 
  filter(!is.na(`Per capita mismanaged plastic waste (kilograms per person per day)`)) %>%
  count(Year)
```


```{r}
waste_vs_gdp %>% 
  filter(!is.na(`Per capita plastic waste (kilograms per person per day)`)) %>%
  count(Year)
```

* There are overlapping variables: `Entity`、`Code`、`Total Population` appears in all 3, with the first two variables allowing for the possibility of joining, and the last being redundant.  


Here we could first define a function to do basic cleaning to all 3 datasets to avoid repetition:  

```{r}
clean_dataset <- function(tbl) {
  tbl %>%
    clean_names() %>%
    filter(year == 2010) %>%
    select(-year) %>%
    rename(country = entity,
           country_code = code)
}
```

Then, join 3 datasets, delete redundant columns and continue renaming (this time using `select` instead of `rename`)

```{r}
coast_vs_waste %>%
  clean_dataset() %>%
  select(-total_population_gapminder) %>%
  inner_join(clean_dataset(mismanaged_vs_gdp) %>%
               select(-total_population_gapminder), by = c("country", "country_code")) %>%
  inner_join(clean_dataset(waste_vs_gdp), by = c("country", "country_code")) %>% 
  select(country, 
         country_code,
         mismanaged_waste = mismanaged_plastic_waste_tonnes,
         coastal_population,
         total_population = total_population_gapminder,
         mismanaged_per_capita = per_capita_mismanaged_plastic_waste_kilograms_per_person_per_day,
         gdp_per_capita = gdp_per_capita_ppp_constant_2011_international_constant_2011_international)
```

Since the [original article](https://ourworldindata.org/faq-on-plastics) focused largely on mismanaged waste and its influence, we'd better filter out any missing values in that variable:  

```{r}
plastic_waste <- coast_vs_waste %>%
  clean_dataset() %>%
  select(-total_population_gapminder) %>%
  inner_join(clean_dataset(mismanaged_vs_gdp) %>%
               select(-total_population_gapminder), by = c("country", "country_code")) %>%
  inner_join(clean_dataset(waste_vs_gdp), by = c("country", "country_code")) %>% 
  select(country, 
         country_code,
         mismanaged_waste = mismanaged_plastic_waste_tonnes,
         coastal_population,
         total_population = total_population_gapminder,
         mismanaged_per_capita = per_capita_mismanaged_plastic_waste_kilograms_per_person_per_day,
         gdp_per_capita = gdp_per_capita_ppp_constant_2011_international_constant_2011_international) %>%
  filter(!is.na(mismanaged_per_capita))
```

